Environment affects specialisation of plants and pollinators

### 2019-12-03 paper/manuscript.Rmd
update paper when bibliography changes
Correct more typos and grammar

loadd(abs_wordcount, msc_wordcount, n_displays, n_references, keywords, bibliography)
loadd(interaction_citations, config) 
```

@@

Species interactions are known to vary widely across space and time. 
There are multiple examples of species that interact with a large number of partners in a particular community or season, but with fewer in another. 
Some of this variation can be attributed to environmental drivers. 
However, how exactly the environment, specifically the stress it imposes on species, affects whether two species interact or not, and ultimately the species' specialisation. 
Understanding how the environment drives the number of partners is crucial because it underpins the species role in its community and shapes the structure of the network of interactions. 
This structure, in turn, determines ecosystem function and stability.

@@

One possible explanation for the seemingly contradictory evidence is that different bioclimatic factors (like temperature or precipitation) can have contrasting effects on species and their partners. 
Here we attempt to simplify this situation by reducing multiple factors into a single measure of environmental stress. 
Previous research suggests that environmental stress may affect the number of partners in different ways depending on its role in the community (for example its trophic guild) or even the species itself. 
Specifically, we propose two alternative hypotheses of how environmental stress may affect specialisation [@tylianakis_ecological_2017]. 
On the one hand, it is possible that when species are under environmental stress, they might be "pressured" to focus on partners with which they are best adapted to interact. 
For instance, @hoiss_altitude_2012 found increased phylogenetic clustering between plants and pollinators at higher altitudes; while @peralta_phylogenetic_2015 found that parasitoids in plantation forest, where environmental stress was higher than in native forests, were constrained to interact with hosts, they were best adapted to attack.
Similarly, @lavandero_genotype_2013 found that environmental stress due to higher temperature reduced the trophic niche breadth of parasitoids suggesting higher specialisation.

@@

Here, we investigate whether and how environmental stress can systematically affect specialisation.  
Our main aim is to test the two hypotheses mentioned above that relate environmental stress and species' number of partners and investigate whether this changes across species or between trophic guilds. 
We propose that specialist species can become "facultative" generalists to reduce their vulnerability to the absence of preferred partners [for example, when variations in climate decouple phenologies; @benadi_specialization_2014]. 
In other words, we expect that, as environmental stress increases, specialists should be more likely to engage with more partners. 
Species with a large number of partners, on the other hand, should have a larger pool of available partners and might, therefore, be more likely to specialise under environmental stress and focus on the most beneficial partners. 
Importantly, when testing these hypotheses, we control for the potential effects of the environment in community composition [which has been previously shown to be a determinant factor; @gravel_bringing_2018]
We test these hypotheses using data on plant-pollinator interactions. 
These interactions provide a particularly interesting system to test these hypotheses because, due to the multiple trade-offs involved in the pollination service, there are multiple intuitive ways in which we could imagine species respond to environmental stress given the available partners.
We estimate the stress species might experience in their community by calculating the bioclimatic suitability of their communities given the species' patterns of global occurrence.

@@

fig_dist_species_multiple_locations_legend <- 
  paste0("Frequency distribution of the number of locations in which a species is present. The most common pollinator species was *", 
       most_popular_species$sp_name[1], 
       "*, which was sampled on ", 
       most_popular_species$n_locations[1], 
       " locations, while the most common plant species was *", 
       most_popular_species$sp_name[2], 
       "*, which was sampled on ", 
       most_popular_species$n_locations[2], 
       " locations.")

@@

### Environmental stress

@@

The niche factor analysis requires two critical pieces of information. 
First, we require information about the occurrences of the species of interest.
Second, we require information about the environmental conditions for all the locations in which the species occurs.
We retrieved `r round(sum(n_occurrences$N)/1000000, digits = 1)` million occurrences from the Global Biodiversity Information Facility (GBIF; https://www.gbif.org). 
Issues with data quality are a central issue hampering the use of publicly available species occurrence GBIF data in ecology and biogeography [@jetz_essential_2019]. 
We, therefore, followed a series of filters and geographic heuristics to correct or remove erroneous and imprecise referencing records [see supplementary methods; @zizka_coordinatecleaner_2019] which allowed us to identify and remove `r round(sum(n_dirty_occurrences$N)/1000000, digits = 1)` million potentially problematic occurrences from further analysis.
We integrated the occurrences from our plant-pollinator communities to the cleaned occurrences retrieved from GBIF.

@@

A common issue of terrestrial bioclimatic datasets is that the boundaries of the cells with information do not precisely match the landmass boundaries. 
The result of this missmatch is that not all environmental variables were available for `r format(complete_climate_cells[1], big.mark = ",")` of the raster cells with occurrences (`r scales::percent(complete_climate_cells[1]/sum(complete_climate_cells), accuracy = 0.1)` of the total). 
As expected, the vast majority of these problematic cells were close to the shore. 
To address this issue, we calculated the average value of environmental variables within an `r config$climate_buffer/1000`km buffer of the centre of the cell where the variable was missing and used it to approximate the value of the variable in that cell. 
Using this procedure, we were able to fill environmental variables for `r scales::percent(percent_filled_round_one, accuracy = 0.1)` of the cells where they were missing. 
To fill the remaining `r format(complete_filled_cells_1[1], big.mark= ",")` cells, we repeated the aforementioned procedure but instead using a `r config$climate_buffer*2/1000`km buffer. 
We removed from further analysis occurrences located within the `r complete_filled_cells_2[1]` cells for which we were unable to fill environmental variables (`r scales::percent(complete_filled_cells_2[1]/sum(complete_filled_cells_2))` of the total).

@@

We evaluated four models to assess the relative importance of suitability. 
A first model, our baseline model, included five variables. 
<!-- fixed variables (population-level predictors) and two random variables (grouping levels). -->
The predictors in the baseline model were the environmental stress, its number of known possible partners in the community, and both the species guild (plant or a pollinator) and its interaction with environmental stress.
We included the number of known possible partners as a predictor in our models as it allows us to control for the effects of the environment on community composition, effectively accounting for species co-occurrence. 
We calculate this metric by determining the number of partners with which the species is known to interact in any other community.
Controlling for the number of potential partners makes our model a particularly stringent test of our environmental-stress hypotheses because this variable could explain a large proportion of variance. 
Often, the potential and the actual number of partners is the same or very close to each other, especially for rare species present only in a few communities.

@@

As expected, we found a strong and positive relationship between the number of possible interactions and the number of realised interactions in the community. 
There was also a large difference on WAIC between the model that included this predictor and that that excluded it. 
This result indicates that the availability of potential partners—this is, community composition—accounts for a large proportion of the variability in species degree. 
Importantly, our findings relating to the variability of the stress-specialisation relationship were qualitatively unchanged, whether we included this variable or not.

The standard deviation (in the parameters scale) of the community intercepts was `r round(sd_loc_id$Estimate, 2)` [`r round(sd_loc_id$Q2.5, 2)`, `r round(sd_loc_id$Q97.5, 2)`] which indicates the importance of the local context when determining specialisation. 
The standard deviation of the species intercept was `r round(sd_org_id$Estimate, 2)` [`r round(sd_org_id$Q2.5, 2)`, `r round(sd_org_id$Q97.5, 2)`], and that of the species' suitability slope was `r round(sd_org_id_suitability$Estimate, 2)` [`r round(sd_org_id_suitability$Q2.5, 2)`, `r round(sd_org_id_suitability$Q97.5, 2)`] (95% credible intervals shown within square brakets).

@@

Similarly, if the patterns we see in our models have also played a role during the evolutionary history of pollination communities, our results also help explain why only a small fraction of plant-pollinator interactions show a strong signature of deep co-evolutionary history [@hutchinson_cophylogenetic_2017].
The increases on the stress that species are predicted to experience due to rapid environmental change might further erode the co-evolutionary history of specialist species.
Communities as a whole might be in a trajectory of even more diffuse co-evolution.
For specialist, at least, the longer-term benefits of being able to interact with multiple partners might be more important than the shorter-term benefits of interacting with partners of high trait matching.

>>>>

loadd(abs_wordcount, msc_wordcount, n_displays, n_references, keywords, bibliography)
loadd(interaction_citations, config) 
loadd(bibliography)
```

@@

Species interactions are known to vary widely across space and time. 
There are multiple examples of species that interact with a large number of partners in a particular community or season, but with fewer in another. 
Some of this variation can be attributed to environmental drivers. 
However, how exactly the environment, specifically the stress it imposes on species, affects whether two species interact or not, and ultimately the species' specialisation. 
Understanding how the environment drives the number of partners is crucial because it underpins the species' role in its community and shapes the structure of the network of interactions. 
This structure, in turn, determines ecosystem function and stability.

@@

One possible explanation for the seemingly contradictory evidence is that different bioclimatic factors (like temperature or precipitation) can have contrasting effects on species and their partners. 
Here we attempt to simplify this situation by reducing multiple factors into a single measure of environmental stress. 
Previous research suggests that environmental stress may affect the number of partners in different ways depending on its role in the community (for example its trophic guild) or even the species itself. 
Specifically, we propose two alternative hypotheses of how environmental stress may affect specialisation [@tylianakis_ecological_2017]. 
On the one hand, it is possible that when species are under environmental stress, they might be "pressured" to focus on partners with which they are best adapted to interact. 
For instance, @hoiss_altitude_2012 found increased phylogenetic clustering between plants and pollinators at higher altitudes; while @peralta_phylogenetic_2015 found that parasitoids in plantation forest, where environmental stress was higher than in native forests, were constrained to interact with hosts, they were best adapted to attack.
Similarly, @lavandero_genotype_2013 showed that environmental stress due to higher temperature reduced the trophic niche breadth of parasitoids suggesting higher specialisation.

@@

Here, we investigate whether and how environmental stress can systematically affect specialisation. 
Our main aim is to test the two hypotheses mentioned above that relate environmental stress and species' number of partners and investigate whether this changes across species or between trophic guilds. 
We propose that specialist species can become "facultative" generalists to reduce their vulnerability to the absence of preferred partners [for example, when variations in climate decouple phenologies; @benadi_specialization_2014]. 
In other words, we expect that, as environmental stress increases, specialists should be more likely to engage with more partners. 
Species with a large number of partners, on the other hand, should have a larger pool of available partners and might, therefore, be more likely to specialise under environmental stress and focus on the most beneficial partners. 
Importantly, when testing these hypotheses, we control for the potential effects of the environment in community composition [which has been previously shown to be a determinant factor; @gravel_bringing_2018]. 
We test these hypotheses using data on plant-pollinator interactions. 
These interactions provide a particularly interesting system to test these hypotheses because, due to the multiple trade-offs involved in the pollination service, there are multiple intuitive ways in which we could imagine species respond to environmental stress given the available partners.
We estimate the stress species might experience in their community by calculating the bioclimatic suitability of their communities given the species' patterns of global occurrence.

@@

fig_dist_species_multiple_locations_legend <- 
  paste0("Frequency distribution of the number of locations in which a species is present. The most common pollinator species was *", 
       most_popular_species$sp_name[1], 
       "*, which was sampled at ", 
       most_popular_species$n_locations[1], 
       " locations, while the most common plant species was *", 
       most_popular_species$sp_name[2], 
       "*, which was sampled at ", 
       most_popular_species$n_locations[2], 
       " locations.")

@@

## Environmental stress

@@

The niche factor analysis requires two critical pieces of information. 
First, it requires information about the occurrences of the species of interest.
Second, the method requires information about the environmental conditions for all the locations in which the species occurs.
We retrieved `r round(sum(n_occurrences$N)/1000000, digits = 1)` million occurrences from the Global Biodiversity Information Facility (GBIF; https://www.gbif.org). 
Issues with data quality are a central issue hampering the use of publicly available species occurrence GBIF data in ecology and biogeography [@jetz_essential_2019]. 
We, therefore, followed a series of filters and geographic heuristics to correct or remove erroneous and imprecise referencing records [see supplementary methods; @zizka_coordinatecleaner_2019] which allowed us to identify and remove `r round(sum(n_dirty_occurrences$N)/1000000, digits = 1)` million potentially problematic occurrences from further analysis.
We integrated the occurrences from our plant-pollinator communities to the cleaned occurrences retrieved from GBIF.

@@

A common issue of terrestrial bioclimatic datasets is that the boundaries of the cells with information do not precisely match the landmass boundaries. 
The result of this missmatch is that not all environmental variables were available for `r format(complete_climate_cells[1], big.mark = ",")` of the raster cells with occurrences (`r scales::percent(complete_climate_cells[1]/sum(complete_climate_cells), accuracy = 0.1)` of the total). 
As expected, the vast majority of these problematic cells were close to the shore. 
To address this issue, we calculated the average value of environmental variables within a `r config$climate_buffer/1000`km buffer of the centre of the cell where the variable was missing and used it to approximate the value of the variable in that cell. 
Using this procedure, we were able to fill environmental variables for `r scales::percent(percent_filled_round_one, accuracy = 0.1)` of the cells where they were missing. 
To fill the remaining `r format(complete_filled_cells_1[1], big.mark= ",")` cells, we repeated the aforementioned procedure but instead using a `r config$climate_buffer*2/1000`km buffer. 
We removed from further analysis occurrences located within the `r complete_filled_cells_2[1]` cells for which we were unable to fill environmental variables (`r scales::percent(complete_filled_cells_2[1]/sum(complete_filled_cells_2), 0.01)` of the total).

@@

We evaluated four models to assess the relative importance of suitability. 
A first model, our baseline model, included five variables. 
<!-- fixed variables (population-level predictors) and two random variables (grouping levels). -->
The predictors in the baseline model were the environmental stress, its number of known possible partners in the community, and both the species guild (plant or a pollinator) and its interaction with environmental stress.
We included the number of known possible partners as a predictor in our models as it allows us to control for the effects of the environment on community composition, effectively accounting for species co-occurrence. 
We calculated this metric by determining the number of partners with which the species is known to interact in any other community.
Controlling for the number of potential partners makes our model a particularly stringent test of our environmental-stress hypotheses because this variable could explain a large proportion of variance. 
Often, the potential and the actual number of partners is the same or very close to each other, especially for rare species present only in a few communities.

@@

As expected, we found a strong and positive relationship between the number of possible interactions and the number of realised interactions in the community. 
There was also a large difference of WAIC between the model that included this predictor and that that excluded it. 
This result indicates that the availability of potential partners—this is, community composition—accounts for a large proportion of the variability in species degree. 
Importantly, our findings relating to the variability of the stress-specialisation relationship were qualitatively unchanged, whether we included this variable or not.

The standard deviation (in the parameters scale) of the community intercepts was `r round(sd_loc_id$Estimate, 2)` [`r round(sd_loc_id$Q2.5, 2)`, `r round(sd_loc_id$Q97.5, 2)`] which indicates the importance of the local context when determining specialisation. 
The standard deviation of the species intercept was `r round(sd_org_id$Estimate, 2)` [`r round(sd_org_id$Q2.5, 2)`, `r round(sd_org_id$Q97.5, 2)`], and that of the species' stress slope was `r round(sd_org_id_suitability$Estimate, 2)` [`r round(sd_org_id_suitability$Q2.5, 2)`, `r round(sd_org_id_suitability$Q97.5, 2)`] (95% credible intervals shown within square brakets).

@@

Similarly, if the patterns we see in our models have also played a role during the evolutionary history of pollination communities, our results also help explain why only a small fraction of plant-pollinator interactions show a strong signature of deep co-evolutionary history [@hutchinson_cophylogenetic_2017].
The increases in the stress that species are predicted to experience due to rapid environmental change might further erode the co-evolutionary history of specialist species.
Communities as a whole might be in a trajectory of even more diffuse co-evolution.
For specialists, at least, the longer-term benefits of being able to interact with multiple partners might be more important than the shorter-term benefits of interacting with partners of high trait matching.
